Realizing temporal graphs from fastest travel times

### 2023-02-01 ICALP23-RealizingTemporalGraphs.tex
Changes 31st Jan

\documentclass[11pt,a4paper]{article}

\usepackage[english]{babel}

\usepackage{fullpage}

\usepackage{amsmath}
\let\proof\relax
\let\endproof\relax

@@

\newtheorem{theorem}{Theorem}
\newtheorem{observation}{Observation}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conjecture}[theorem]{Conjecture}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[definition]{Example}
\usepackage{comment}

@@

\newenvironment{claimproof}{\noindent\emph{Proof of Claim.}}{\hfill$\diamond$

}

@@

\usepackage[capitalise,nameinlink, noabbrev]{cleveref}
\crefname{claim}{Claim}{Claims}
\crefname{observation}{Observation}{Observations}

\usepackage{authblk}

@@

\newcommand{\ie}{i.\,e.,\ }
\newcommand{\st}{s.\,t.,\ }
\newcommand{\NP}{\textrm{NP}}
\newcommand{\APX}{\textrm{APX}}
\newcommand{\FPT}{\textrm{FPT}}
\newcommand{\XP}{\textrm{XP}}

@@

\title{Realizing Temporal Graphs}
\author{Nina Klobas, George B. Mertzios, Hendrik Molter, Paul G. Spirakis}

\bibliographystyle{abbrv}
\begin{document}
	\maketitle

@@

@@

\section{
\texorpdfstring{\deltaExact\ } {Simple Delta-TGR}}

@@

\subsection{
\texorpdfstring{\deltaExact\ } {Simple Delta-TGR}
is polynomial-time solvable on trees and paths}

@@

\subsection{
\texorpdfstring{\deltaExact\ } {Simple Delta-TGR}
is polynomial-time solvable on cycles}

@@

    $d(c,x_F)=2$.

@@

    $d(c,x_T)=2$.
    \item Let $x$ be a variable that does \emph{not} appear in clause $c$, then we specify the following distances:

@@

All edges incident with $v$ have labels two or three. All edges incident with $x$ have label two and all edges incident with $c$ have label one. We immediately have that there is no temporal path from $v$ to $c$. Furthermore, since $x$ and $v$ are non-adjacent, we also have that there is no temporal path from $v$ to $x$.
\end{itemize}
\end{proof}

\section{W[1]-hardness}
\begin{theorem}\label{thm:W1wrtFVS}
    \deltaExactLong\ is W[1]-hard parameterized by the feedback vertex number.
\end{theorem}
\begin{proof}
    We present a parameterized reduction from the W[1]-hard problem \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval}.  Here, given a $k$-partite graph $H=(W_1\uplus W_2 \uplus\ldots\uplus W_k, F)$, we are asked whether $H$ contains a clique of size $k$. If $w\in W_i$, then we say that $w$ has \emph{color} $i$. W.l.o.g.\ we assume that $|W_1|=|W_2|=\ldots=|W_k|=n$ and that every vertex has at least one neighbor of every color. 
    Furthermore, for all $i\in[k]$, we assume the vertices in $W_i$ are ordered in some arbitrary but fixed way, that is, $W_i=\{w^i_1,w^i_2,\ldots,w^i_n\}$.
    Let $F_{i,j}$ with $i<j$ denote the set of all edges between vertices from $W_i$ and $W_j$. We assume w.l.o.g.\ that $|F_{i,j}|=m$ for all $i< j$.
    Furthermore, for all $i<j$ we assume that the edges in $F_{i,j}$ are ordered in some arbitrary but fixed way.

@@

    We will construct $G$ using several gadgets. We first introduce an \emph{edge selection gadget} for color combination $i,j$ with $i<j$. We start with describing the vertex set of the gadget.
    \begin{itemize}
        \item A set $X_{i,j}$ of vertices $x_1, x_2, \ldots, x_m$.
        \item A set $Y_{i,j}$ of vertices $y_1, y_2, \ldots, y_m$.
        \item A set $Z_{i,j}$ of vertices $z_1, z_2, \ldots, z_m$.
        \item Vertex sets $U_1, U_2, \ldots, U_m$ with $U_\ell=\{u^\ell_1, u^\ell_2,\ldots, u^\ell_n, u^\ell_{n+1}, u^\ell_{n+2}, u^v_{2n}\}$ for all $\ell\in[m]$. Let $U_{i,j}=\bigcup_{\ell\in[n]} U_\ell$.
        \item Two special vertices $v_{i,j}^\star,v_{i,j}^{\star\star}$.
    \end{itemize}
    The gadget has the following edges.
    \begin{itemize}
        \item For all $\ell\in [m]$ we have edge $\{x_\ell,v_{i,j}^\star\}$, $\{y_\ell,v_{i,j}^\star\}$, and $\{z_\ell,v_{i,j}^{\star\star}\}$.

        Furthermore, we have edges $\{y_\ell,u^\ell_1\}$ and $\{z_\ell,u^\ell_{2n}\}$.
        \item For all $\ell\in [m]$ and $\ell'\in [2n-1]$, we have edge $\{u^\ell_{\ell'},u^\ell_{\ell'+1}\}$.
    \end{itemize}
    This finishes the construction of the edge selection gadget $G_{i,j}$.

@@

    Let $1\le j<i$ (skip if $i=1$), let $e_\ell\in F_{j,i}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v_{j-1}^i,u^\ell_{\ell'}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $j,i$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{\ell'+1}\}$ between $v^i_j$ and the vertex $u^\ell_{\ell'+1}$ of the edge selection gadget of color combination $j,i$.

    If $1\neq i\neq k$, then we add edge $\{v^i_{i-1},\hat{u}^i_1\}$ and for all $\ell\in[n+2]$ we add edge $\{\hat{u}^i_\ell,\hat{u}^i_{\ell+1}\}$. 
    Let $e_\ell\in F_{i,i+1}$ and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. If $i\neq 1\neq k$, then we add edge $\{\hat{u}^i_{n+3},u^\ell_{n+\ell'}\}$ between $\hat{u}^i_{n+3}$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,i+1$. If $i=1$, then we add edge $\{v^i_0,u^\ell_{n+\ell'}\}$ between $v^i_0$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,i+1$.

    Let $i<j\le k$ (skip if $i=k$), let $e_\ell\in F_{i,j}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell$. Then we add edge $\{v_{j-1}^i,u^\ell_{n+\ell'}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,j$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{n+\ell'+1}\}$ between $v^i_j$ and the vertex $u^\ell_{n+\ell'+1}$ of the edge selection gadget of color combination $i,j$.

    Next, we describe \emph{connector gadgets}. Intuitively, these gadgets will be used to connect many vertex pairs by short paths, which will make arguing about possible labelings in yes-instances much easier. Connector gadgets consist of three vertices $\hat{v}_0,\hat{v}_1,\hat{v}_2$. 
    %Each connector gadget is associated with four sets $A,A',B,B'$ containing vertices of other gadgets. 
    Each connector gadget is associated with two sets $A,B$ with $B\subseteq A$ containing vertices of other gadgets. 
    Let $V^\star$ denote the set of all vertices from all edge selection gadgets and all validation gadgets.
    The sets $A$ and $B$ will only play a role when defining the distance matrix $D$ later. Informally speaking, vertices in $A$ should reach all vertices in $V^\star$ quickly through the gadget, except the ones in $B$.
    We have the following edges. 
    \begin{itemize}
        \item Edges $\{\hat{v}_0,\hat{v}_1\},\{\hat{v}_1,\hat{v}_2\}$.
        \item An edge between $\hat{v}_1$ and each vertex in $V^\star$.
        \item An edge between $\hat{v}_2$ and each vertex in $V^\star$.
    \end{itemize}
    We add two connector gadgets for each edge selection gadget and one connector gadget for each validation gadget.

@@

Next, consider a connector gadget consisting of vertices $\hat{v}_0,\hat{v}_1,\hat{v}_2$ and with sets $A$ and $B$. Informally, the connector gadget makes sure that all vertices in $A$ can reach all other vertices (of edge selection gadgets and validation gadgets) except the ones in $B$. We set the following distances. Recall that $V^\star$ denotes the set of all vertices from all edge selection gadgets and all validation gadgets.
\begin{itemize}
    \item We set $d(\hat{v}_0,\hat{v}_2)=2$.
    \item Let $v\in V^\star\setminus B$, then we set $d(\hat{v}_0,v)=3$.
    \item Let $v\in A$ and $v'\in V^\star\setminus B$ such that $v$ and $v'$ are not neighbors, then we set $d(v,v')=3$.
\end{itemize}
Now consider two connector gadgets, one with vertices $\hat{v}_0,\hat{v}_1,\hat{v}_2$ and with sets $A$ and $B$, and one with vertices $\hat{v}_0',\hat{v}_1',\hat{v}_2'$ and with sets $A'$ and $B'$.
\begin{itemize}
    \item If there is a vertex $v\in A$ with $v\notin A'$, then we set $d(\hat{v}_1,\hat{v}_1')=3$.
    \item If there is a vertex $v\in A$ with $v\in A'\setminus B'$, then we set $d(\hat{v}_1,\hat{v}_2')=3$.
    \item If there is a vertex $v\in V^\star\setminus (A\setminus B)$ with $v\notin A'$, then we set $d(\hat{v}_2,\hat{v}_1')=3$.    
    \item If there is a vertex $v\in V^\star\setminus (A\setminus B)$ with $v\in A'\setminus B'$, then we set $d(\hat{v}_2,\hat{v}_2')=3$.
\end{itemize}

Next, consider the edge selection gadget for color combination $i,j$ with $i<j$.
\begin{itemize}
    \item Let $1\le \ell<\ell'\le m$. We set $d(x_\ell,x_\ell')=2n\cdot (i+j)\cdot(\ell'-\ell)+1$.
    \item For all $\ell\in[m]$ we set $d(x_\ell,v_{i,j}^{\star\star})=4n+3$.
\end{itemize}

>>>>

\documentclass[a4paper,UKenglish,cleveref, autoref, thm-restate,anonymous]{lipics-v2021}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository

%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory

\usepackage{amsmath}
%\let\proof\relax
%\let\endproof\relax

@@

\usepackage{comment}

@@

@@

\crefname{claim}{Claim}{Claims}

@@

\newcommand{\ie}{i.\,e.,\ }
\renewcommand{\st}{s.\,t.,\ }
\newcommand{\NP}{\textrm{NP}}
\newcommand{\APX}{\textrm{APX}}
\newcommand{\FPT}{\textrm{FPT}}
\newcommand{\XP}{\textrm{XP}}

@@

\bibliographystyle{plainurl}% the mandatory bibstyle

\title{Realizing temporal graphs from fastest travel times} %TODO Please add

%\titlerunning{Dummy short title} %TODO optional, please use if title is longer than one line

%\author{Nina Klobas}{Department of Computer Science, Durham University, UK}{nina.klobas@durham.ac.uk}{ https://orcid.org/0000-0002-8024-5782}{}

%\author{George B. Mertzios}{Department of Computer Science, Durham University, UK}{george.mertzios@durham.ac.uk}{https://orcid.org/0000-0001-7182-585X}{Supported by the EPSRC grant EP/P020372/1.}

%\author{Hendrik Molter}{Department of Industrial Engineering and Management, Ben-Gurion~University~of~the~Negev, Israel}{molterh@post.bgu.ac.il}{https://orcid.org/0000-0002-4590-798X}{Supported by the ISF, grant No.~1070/20.}

%\author{Paul G. Spirakis}{Department of Computer Science, University of Liverpool, UK}{p.spirakis@liverpool.ac.uk}{https://orcid.org/0000-0001-5396-3749}{Supported by the EPSRC grant EP/P02002X/1.}

%\authorrunning{Nina Klobas, George B. Mertzios, Hendrik Molter, and Paul G. Spirakis} %TODO mandatory. First: Use abbreviated first/middle names. Second (only in severe cases): Use first author plus 'et al.'

%\Copyright{Nina Klobas, George B. Mertzios, Hendrik Molter, and Paul G. Spirakis} %TODO mandatory, please use full first names. LIPIcs license is "CC-BY";  http://creativecommons.org/licenses/by/3.0/

\ccsdesc[500]{Theory of computation~Graph algorithms analysis}
\ccsdesc[500]{Mathematics of computing~Discrete mathematics}
%TODO mandatory: Please choose ACM 2012 classifications from https://dl.acm.org/ccs/ccs_flat.cfm

\keywords{Temporal graph, periodic temporal labeling, fastest temporal path, graph realization, temporal connectivity.} %TODO mandatory; please add comma-separated list of keywords

\category{} %optional, e.g. invited paper

\relatedversion{} %optional, e.g. full version hosted on arXiv, HAL, or other respository/website
%\relatedversiondetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93]{Classification (e.g. Full Version, Extended Version, Previous Version}{URL to related version} %linktext and cite are optional

%\supplement{}%optional, e.g. related research data, source code, ... hosted on a repository like zenodo, figshare, GitHub, ...
%\supplementdetails[linktext={opt. text shown instead of the URL}, cite=DBLP:books/mk/GrayR93, subcategory={Description, Subcategory}, swhid={Software Heritage Identifier}]{General Classification (e.g. Software, Dataset, Model, ...)}{URL to related version} %linktext, cite, and subcategory are optional

%\funding{(Optional) general funding statement \dots}%optional, to capture a funding statement, which applies to all authors. Please enter author specific funding statements as fifth argument of the \author macro.

%\acknowledgements{I want to thank \dots}%optional

%\nolinenumbers %uncomment to disable line numbering

%Editor-only macros:: begin (do not touch as author)%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\EventEditors{John Q. Open and Joan R. Access}
\EventNoEds{2}
\EventLongTitle{42nd Conference on Very Important Topics (CVIT 2016)}
\EventShortTitle{CVIT 2016}
\EventAcronym{CVIT}
\EventYear{2016}
\EventDate{December 24--27, 2016}
\EventLocation{Little Whinging, United Kingdom}
\EventLogo{}
\SeriesVolume{42}
\ArticleNo{23}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\maketitle

\begin{abstract}
In this paper we study \emph{temporal design} problems of undirected temporally connected graphs. 
The basic setting of these optimization problems is as follows: given an undirected graph~$G$, 
what is the smallest number $|\lambda|$ of time-labels that we need to add to the edges of $G$ such that the resulting temporal graph $(G,\lambda)$ is temporally connected?

\vspace{0,2cm}
\noindent \textbf{Due to lack of space, the full paper with all proofs is attached in a 
	clearly marked Appendix to be read at the discretion of the Program Committee.}
\end{abstract}

\section{Introduction}\label{intro-sec}

Graph realization problems are fundamental graph-algorithmic questions where the objective is to (re)construct a graph, 
given some constraints on some of its graph attributes. Among the most well studied such parameters for graph reconstruction 
are the vertex degrees~\cite{}, distances~\cite{}, eccentricities~\cite{}, connectivity~\cite{}, etc. 
The motivation for network realization problems stems from both ``verification'' applications where we attempt to reconstruct an existing network of unknown structure based on the outcomes of experimental measurements, and engineering applications related to network design~\cite{.....}.

@@

\section{Motivation}

	When considering the problem of finding shortest paths among vertices, one relies heavily on the optimal substructure of the problem,
	which implies that if a shortest path $P$ from vertex $u$ to $v$ traverses the vertex $w$,
	then the shortest path from $u$ to $w$ is a subpath of $P$.
	After studying the problem of fastest temporal paths (\ie temporal paths of smallest duration),
	one can see that this property does not carry over.
	For an example see~\cref{fig:ftpExample}.
	Which makes all the standard intuition and work with these kind of paths, much harder.

	\begin{figure}[h]
    \centering
	\begin{tikzpicture}[xscale=2]
	%%%S_uv
	\node[vert,label=below:$u$] (v1) at (1,0) {};
	\node[vert] (v2) at (2,0) {};
	\node[vert,label=below:$w$] (v3) at (3,0) {};
	\node[vert] (v4) at (4,0) {};
	\node[vert,label=below:$v$] (v5) at (5,0) {};
	\draw (v1) -- node[timelabel] {$1$} (v2) -- node[timelabel] {$3$}  (v3) -- node[timelabel] {$5$} (v4) -- node[timelabel] {$7$} (v5);
	\draw (v1) to [out=80,in=100,distance=2cm] node[timelabel] {$9$} (v3);

	\draw[transform canvas={yshift=-1.5mm}, blue]
	(v1) edge[diredge2]  node[pos=0.3,yshift=-2,label=below:$P_{u,v}$] {} (v5) ;

	\draw[red]
	(v1) edge[diredge2] [out=85,in=95,distance=2.1cm] node[pos=0.3,yshift=2,label=above:$P_{u,w}$] {} (v3) ;
\end{tikzpicture}
	\caption{An example of a temporal graph, where the fastest temporal path $P_{u,v}$ (in blue) from $u$ to $v$ is of duration $7$,
		while the fastest temporal path $P_{u,w}$ (in red) from $u$ to a vertex $w$ that is on a path $P_{u,v}$ is of duration $1$ and is not a subpath of $P_{u,v}$.		
		\label{fig:ftpExample}}
	\end{figure}

@@

\section{\deltaExact}

@@

\subsection{ \deltaExact\ is polynomial-time solvable on trees and paths}

@@

\subsection{ \deltaExact\
is polynomial-time solvable on cycles}

@@

    $d(c,x^F)=2$.

@@

    $d(c,x^T)=2$.
    \item Let $x$ be a variable that does \emph{not} appear in clause $c$, then we specify the following distances:

@@

All edges incident with $v$ have labels two or three. All edges incident with $x$ have label two and all edges incident with $c$ have label one. We immediately have that there is no temporal path from $v$ to $c$. Furthermore, since $x$ and $v$ are non-adjacent, we also have that there is no temporal path from $v$ to $x$.
\end{itemize}
%
\begin{figure}[h]
	\noindent
	\makebox[\textwidth]{
		\centering
\begin{tikzpicture}%[xscale=2]

	%%Variable X1
	\node[vert,label=left:$x_1^T$] (x1t) at (1,0) {};
	\node[vert,label=right:$x_1^F$] (x1f) at (3,0) {};
	\node[vert,label=above:$x_1$] (x1) at (2,1) {};
	\draw (x1t) -- (x1f) -- (x1) -- (x1t);

	%%Variable X2
	\node[vert,label=left:$x_2^T$] (x2t) at (5,0) {};
	\node[vert,label=right:$x_2^F$] (x2f) at (7,0) {};
	\node[vert,label=above:$x_2$] (x2) at (6,1) {};
	\draw (x2t) -- (x2f) -- (x2) -- (x2t);

	%%Variable x3
	\node[vert,label=left:$x_3^T$] (x3t) at (9,0) {};
	\node[vert,label=right:$x_3^F$] (x3f) at (11,0) {};
	\node[vert,label=above:$x_3$] (x3) at (10,1) {};
	\draw (x3t) -- (x3f) -- (x3) -- (x3t);

	%%Variable x4
	\node[vert,label=left:$x_4^T$] (x4t) at (13,0) {};
	\node[vert,label=right:$x_4^F$] (x4f) at (15,0) {};
	\node[vert,label=above:$x_4$] (x4) at (14,1) {};
	\draw (x4t) -- (x4f) -- (x4) -- (x4t);

	%%vertices for clauses c1, c2 + vertex v
	\node[vert,label=above:$v$] (v) at (8,3) {};
	\node[vert,label=below:$c_1$] (c1) at (4.5,-3) {};
	\node[vert,label=below:$c_2$] (c2) at (11.5,-3) {};

	%%% edges between x1 gadget and v,c1,c2
	\draw (x1t) to node[timelabel] {$1$} (c1);
	\draw (x1t) to node[timelabel] {$1$} (c2);
	\draw (x1t) to [out=90,in=190] (v);
	\draw (x1f) to [out=90,in=200] node[timelabel] {$3$} (v);

	%%% edges between x2 gadget and v,c1,c2
	\draw (x2f) to node[timelabel] {$1$} (c1);
	\draw (x2t) to node[timelabel] {$1$} (c2);
	\draw (x2t) to [out=80,in=210]  (v);
	\draw (x2f) to node[timelabel] {$3$} (v);

	%%% edges between x3 gadget and v,c1
	\draw (x3t) to node[timelabel] {$1$} (c1);
	\draw (x3t) to node[timelabel] {$3$} (v);
	\draw (x3f) to [out=100,in=330] (v);

	%%% edges between x4 gadget and v,c2
	\draw (x4t) to node[timelabel] {$1$} (c2);
	\draw (x4t) to [out=90,in=340] node[timelabel] {$3$} (v);
	\draw (x4f) to [out=90,in=360] (v);

	%% edges among xt, xf vertices
	%x1 to all
	\draw [dotted] (x1t) [out=30,in=150] to (x2t);
	\draw [dotted] (x1t) [out=30,in=150] to (x3t);
	\draw [dotted] (x1t) [out=30,in=150] to (x4t);
	\draw [dotted] (x1t) [out=30,in=150] to (x2f);
	\draw [dotted] (x1t) [out=30,in=150] to (x3f);
	\draw [dotted] (x1t) [out=30,in=150] to (x4f);

	\draw [dotted] (x1f) [out=30,in=150] to (x2t);
	\draw [dotted] (x1f) [out=30,in=150] to (x3t);
	\draw [dotted] (x1f) [out=30,in=150] to (x4t);
	\draw [dotted] (x1f) [out=30,in=150] to (x2f);
	\draw [dotted] (x1f) [out=30,in=150] to (x3f);
	\draw [dotted] (x1f) [out=30,in=150] to (x4f);
	%x2 to all
	\draw [dotted] (x2t) [out=30,in=150] to (x3t);
	\draw [dotted] (x2t) [out=30,in=150] to (x4t);
	\draw [dotted] (x2t) [out=30,in=150] to (x3f);
	\draw [dotted] (x2t) [out=30,in=150] to (x4f);

	\draw [dotted] (x2f) [out=30,in=150] to (x3t);
	\draw [dotted] (x2f) [out=30,in=150] to (x4t);
	\draw [dotted] (x2f) [out=30,in=150] to (x3f);
	\draw [dotted] (x2f) [out=30,in=150] to (x4f);
	%x3 to all
	\draw [dotted] (x3t) [out=30,in=150] to (x4t);
	\draw [dotted] (x3t) [out=30,in=150] to (x4f);

	\draw [dotted] (x3f) [out=30,in=150] to (x4t);
	\draw [dotted] (x3f) [out=30,in=150] to (x4f);
\end{tikzpicture}
}
\caption{Example of the temporal graph $(G_\phi,\lambda)$ from the NP-hardness reduction, 
	where the 3-SAT formula $\phi$ is of form $\phi = (x_1 \vee \overline{x_2} \vee x_3) \wedge (\overline{x_1} \vee x_2 \vee x_4)$.
For the simplicity we draw edges between vertices $x_i^T$ and $x_j^F$ (where $i \neg j$) with a dotted line.
Presented is the labeling of $G_\phi$, corresponding to the assignment $x_1=x_2=\textsc{True}$ and $x_3,x_4=\textsc{False}$,
where all unlabeled edges get the label $2$.}
\end{figure}
\end{proof}

\section{W[1]-hardness}
\begin{theorem}\label{thm:W1wrtFVS}
    \deltaExactLong\ is W[1]-hard parameterized by the feedback vertex number.
\end{theorem}
\begin{proof}
    We present a parameterized reduction from the W[1]-hard problem \textsc{Multicolored Clique} parameterized by the number of colors~\cite{fellows2009multipleinterval}.  Here, given a $k$-partite graph $H=(W_1\uplus W_2 \uplus\ldots\uplus W_k, F)$, we are asked whether $H$ contains a clique of size $k$. If $w\in W_i$, then we say that $w$ has \emph{color} $i$. W.l.o.g.\ we assume that $|W_1|=|W_2|=\ldots=|W_k|=n$ and that every vertex has at least one neighbor of every color. 
    Furthermore, for all $i\in[k]$, we assume the vertices in $W_i$ are ordered in some arbitrary but fixed way, that is, $W_i=\{w^i_1,w^i_2,\ldots,w^i_n\}$.
    Let $F_{i,j}$ with $i<j$ denote the set of all edges between vertices from $W_i$ and $W_j$. We assume w.l.o.g.\ that $|F_{i,j}|=m$ for all $i< j$.
    Furthermore, for all $i<j$ we assume that the edges in $F_{i,j}$ are ordered in some arbitrary but fixed way, that is, $F_{i,j}=\{e^{i,j}_1,e^{i,j}_2,\ldots,e^{i,j}_m\}$.

@@

    We will construct $G$ using several gadgets. We first introduce an \emph{edge selection gadget} for color combination $i,j$ with $i<j$. We start with describing the vertex set of the gadget.
    \begin{itemize}
        \item A set $X_{i,j}$ of vertices $x_1, x_2, \ldots, x_m$.
        %\item A set $Y_{i,j}$ of vertices $y_1, y_2, \ldots, y_m$.
        %\item A set $Z_{i,j}$ of vertices $z_1, z_2, \ldots, z_m$.
        \item Vertex sets $U_1, U_2, \ldots, U_m$ with $U_\ell=\{u^\ell_0, u^\ell_1, u^\ell_2,\ldots, u^\ell_n, u^\ell_{n+1}, u^\ell_{n+2}, u^\ell_{2n}\}$ for all $\ell\in[m]$. Let $U_{i,j}=\bigcup_{\ell\in[n]} U_\ell$.
        \item Two special vertices $v_{i,j}^\star,v_{i,j}^{\star\star}$.
    \end{itemize}
    The gadget has the following edges.
    \begin{itemize}
        \item For all $\ell\in [m]$ we have edge $\{x_\ell,v_{i,j}^\star\}$, $\{v_{i,j}^\star,u^\ell_0\}$, and $\{u^\ell_{2n},v_{i,j}^{\star\star}\}$.

        %Furthermore, we have edges $\{y_\ell,u^\ell_1\}$ and $\{z_\ell,u^\ell_{2n}\}$.
        \item For all $\ell\in [m]$ and $\ell'\in [2n]$, we have edge $\{u^\ell_{\ell'-1},u^\ell_{\ell'}\}$.
    \end{itemize}
    This finishes the construction of the edge selection gadget $G_{i,j}$.

@@

    Let $1\le j<i$ (skip if $i=1$), let $e_\ell^{j,i}\in F_{j,i}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell^{j,i}$. Then we add edge $\{v_{j-1}^i,u^\ell_{\ell'-1}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{\ell'-1}$ of the edge selection gadget of color combination $j,i$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{\ell'}\}$ between $v^i_j$ and the vertex $u^\ell_{\ell'}$ of the edge selection gadget of color combination $j,i$.

    If $1\neq i\neq k$, then we add edge $\{v^i_{i-1},\hat{u}^i_1\}$ and for all $\ell\in[n+2]$ we add edge $\{\hat{u}^i_\ell,\hat{u}^i_{\ell+1}\}$. 
    Let $e_\ell^{i,i+1}\in F_{i,i+1}$ and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell^{i,i+1}$. If $i\neq 1\neq k$, then we add edge $\{\hat{u}^i_{n+3},u^\ell_{n+\ell'-1}\}$ between $\hat{u}^i_{n+3}$ and the vertex $u^\ell_{n+\ell'-1}$ of the edge selection gadget of color combination $i,i+1$. If $i=1$, then we add edge $\{v^i_0,u^\ell_{n+\ell'-1}\}$ between $v^i_0$ and the vertex $u^\ell_{n+\ell'-1}$ of the edge selection gadget of color combination $i,i+1$.

    Let $i<j\le k$ (skip if $i=k$), let $e_\ell^{i,j}\in F_{i,j}$, and let $w^i_{\ell'}\in W_i$ be incident with $e_\ell^{i,j}$. Then we add edge $\{v_{j-1}^i,u^\ell_{n+\ell'-1}\}$ between $v^i_{j-1}$ and the vertex $u^\ell_{n+\ell'-1}$ of the edge selection gadget of color combination $i,j$.
    Furthermore, we add edge $\{v_{j}^i,u^\ell_{n+\ell'}\}$ between $v^i_j$ and the vertex $u^\ell_{n+\ell'}$ of the edge selection gadget of color combination $i,j$.

    Next, we describe \emph{connector gadgets}. Intuitively, these gadgets will be used to connect many vertex pairs by short paths, which will make arguing about possible labelings in yes-instances much easier. Connector gadgets consist of six vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$. 
    %Each connector gadget is associated with four sets $A,A',B,B'$ containing vertices of other gadgets. 
    Each connector gadget is associated with two sets $A,B$ with $B\subseteq A$ containing vertices of other gadgets. 
    Let $V^\star$ denote the set of all vertices from all edge selection gadgets and all validation gadgets.
    The sets $A$ and $B$ will only play a role when defining the distance matrix $D$ later. Informally speaking, vertices in $A$ should reach all vertices in $V^\star$ quickly through the gadget, except the ones in $B$.
    We have the following edges. 
    \begin{itemize}
        \item Edges $\{\hat{v}_0,\hat{v}_1\},\{\hat{v}_0',\hat{v}_1\},\{\hat{v}_1,\hat{v}_2\},\{\hat{v}_2,\hat{v}_3\},\{\hat{v}_2,\hat{v}_3'\}$.
        \item An edge between $\hat{v}_1$ and each vertex in $V^\star$.
        \item An edge between $\hat{v}_2$ and each vertex in $V^\star$.
    \end{itemize}
    We add two connector gadgets for each edge selection gadget and one connector gadget for each validation gadget.

@@

Next, consider a connector gadget consisting of vertices $\hat{v}_0,\hat{v}_1,\hat{v}_2,\hat{v}_3$ and with sets $A$ and $B$. Informally, the connector gadget makes sure that all vertices in $A$ can reach all other vertices (of edge selection gadgets and validation gadgets) except the ones in $B$. We set the following distances. Recall that $V^\star$ denotes the set of all vertices from all edge selection gadgets and all validation gadgets.
\begin{itemize}
    \item We set $d(\hat{v}_0,\hat{v}_2)=d(\hat{v}_3,\hat{v}_1)=d(\hat{v}_2,\hat{v}_0')=d(\hat{v}_1,\hat{v}_3')=2$, and $d(\hat{v}_0,\hat{v}_0')=d(\hat{v}_3,\hat{v}_3')=d(\hat{v}_0,\hat{v}_3')=d(\hat{v}_3,\hat{v}_0')=3$.
    \item Let $v\in A$, then we set $d(v,\hat{v}_0')=3$ and $d(v,\hat{v}_3')=3$.
    \item Let $v\in V^\star\setminus B$, then we set $d(\hat{v}_0,v)=3$ and $d(\hat{v}_3,v)=3$.
    \item Let $v\in A$ and $v'\in V^\star\setminus B$ such that $v$ and $v'$ are not neighbors, then we set $d(v,v')=3$.
\end{itemize}
Now consider two connector gadgets, one with vertices $\hat{v}_0,\hat{v}_0',\hat{v}_1,\hat{v}_2,\hat{v}_3,\hat{v}_3'$ and with sets $A$ and $B$, and one with vertices $\hat{v}_0',\hat{v}_0'',\hat{v}_1',\hat{v}_2',\hat{v}_3',\hat{v}_3''$ and with sets $A'$ and $B'$.
\begin{itemize}
    \item If there is a vertex $v\in A$ with $v\notin A'$, then we set $d(\hat{v}_1,\hat{v}_1')=3$.
    \item If there is a vertex $v\in A$ with $v\in A'\setminus B'$, then we set $d(\hat{v}_1,\hat{v}_2')=3$.
    \item If there is a vertex $v\in V^\star\setminus (A\setminus B)$ with $v\notin A'$, then we set $d(\hat{v}_2,\hat{v}_1')=3$.    
    \item If there is a vertex $v\in V^\star\setminus (A\setminus B)$ with $v\in A'\setminus B'$, then we set $d(\hat{v}_2,\hat{v}_2')=3$.
\end{itemize}

Next, consider the edge selection gadget for color combination $i,j$ with $i<j$.
\begin{itemize}
    \item Let $1\le \ell<\ell'\le m$. We set $d(x_\ell,x_{\ell'})=2n\cdot (i+j)\cdot(\ell'-\ell)+1$.
    \item For all $\ell\in[m]$ we set $d(x_\ell,v_{i,j}^{\star\star})=4n+3$.
\end{itemize}
